---
title: "Homework guideline"
author: "Your name here"
date: "Due date here"
output:
  html_document
---

```{r global_options, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

***General instructions for homeworks***: 

+  Make a new R Markdown file (.Rmd) referring to the assignment on the course Github page
+  Change the heading to include the homework number, your name, and the due date
+  Save the R Markdown file (named as:  [MikeID]-[hw01].Rmd -- e.g. "mlopez-hw01.Rmd") to somewhere where you'll be able to access it later (zip drive, My Documents, Dropbox, etc)
+  Your file should contain the code/commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file
+  **Each answer must be supported by written statements (unless otherwise specified) as well as any code used**: In other words, if the answer is `24`, you should write "The answer is 24" (as opposed to just showing the code and output).
+  Include the names of anyone you collaborated with at the top of the assignment
+  I recommend copying the raw .Rmd code from the Github page (link [https://raw.githubusercontent.com/statsbylopez/FYE_18/master/HWs/HW%20base.Rmd](https://raw.githubusercontent.com/statsbylopez/FYE_18/master/HWs/HW%20base.Rmd). 
+  Homeworks are due at the start of class -- please print the HTML and hand in. Be sure to staple your homework and ensure that it is in an easily readable format. 


1. *Print* and *Read* Chapters 1 through 4 in R for Data Science, which can be found at ([http://r4ds.had.co.nz/](http://r4ds.had.co.nz/)). You do not need to show code or output for this answer. 

3. Identify what is wrong with the following code chunks:

-  

```{r}
Prof.lopez <- "awesome"
Prof.1opez
```

-  

```{r}
mtcars %>% mutate(gear.times.carb = gear(carb))
```

-  

```{r}
mtcars %>% filter(cyy == 6)
```


-  

```{r}
mtcars %>% filter(vs = 0)
```


4. One of the more useful data sets in R is `nycflights13`. Download this package, and ensure that you have it loaded in R. Within that package is a data set called `flights`, which you can take a look at using the following code.

- How many rows (flights) and columns (variables) are there in the flight data set?

- Identify a numerical variable and a character variable in the `flights` data. 

- Identify the average departure delay (`dep_delay`) in the flights data set.

- Use a tool that we have not yet discussed in class to answer **any** question of interest. As an example, what day of the month do most flights leave on? Do not answer this question, but answer something similar. I encourage you to use outside sources for code!

- Using Chapter 3 of the textbook, show *two* plots using the `flights` data set. They can be any type of plot -- try a few kinds!

- Add better x-axis labels, y-axis labels, a title, and a subtitle to each of your two plots. Note: you only need to show the final versions of the two plots. 


