---
title: "Homework or lab title here"
author: "Your name here"
date: "Date here"
output:
  html_document
---

```{r global_options, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
```



## Questions

This lab will be mostly review, but done on a hockey data set. Please see our previous labs and homeworks for reminders of how to answer the following questions. The answers are more open ended -- be concise in your answers, but feel free to try different plots or summary statistics to justify your answer. 

```{r}
library(RCurl)
library(tidyverse)
url <- getURL("https://raw.githubusercontent.com/statsbylopez/StatsSports/master/NHL.csv")
nhl.data <- read_csv(url)
nhl.data <- nhl.data %>% filter(TOI > 500)
nhl.data <- na.omit(nhl.data)
nhl.data$ShP <- nhl.data$Goals/nhl.data$Shots
```


## Question 1

A player is curious if salaries have changed over time. What will you tell the player? Use a combination of graphs (think about which chart(s) are most appropriate) and summary statistics (mean, median, etc) to identify how salaries have changed. 

## Question 2

Refer to homework 2 (as well as the corresponding code in our notes). Make a correlation matrix featuring the variables `Goals_Sixty`, `Assists_Sixty`, `CF_Percent`, `PDO`, `Shots`, `TOI`, and `SHP`. Which variables are most closely related to one another? 

## Question 3

Identify the 4 player-seasons with the highest PDO. Next, filtering each of those players' one at a time, identify their PDO in the season following their season with a high PDO. 

## Question 4

Using players who have played at least 5 seasons (recall: each row is a season), identify the five players with the highest average `CF_Percent`. The top player should be Pavel.Datsyuk: interpret his average `CF_Percent`. 

## Question 5

Please submit a project proposal, per guidelines discussed in class. The proposal should be typed using RMarkdown and printed as a separate file from this homework. One proposal should be submitted per group.

