---
title: "Homework & lab guidelines"
author: "Your name here"
date: "Date here"
output:
  html_document
---

```{r global_options, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

**General instructions for homeworks and labs**: 

+  Make a new R Markdown file (.Rmd) referring to the assignment on the course Github page
+  Change the heading to include the homework or lab name and the due date
+  Save the R Markdown file (named as:  [MikeID]-[hw01].Rmd -- e.g. "mlopez-hw01.Rmd").
+  Your file should contain the code/commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file
+  **Each answer must be supported by written statements (unless otherwise specified) as well as any code used**: In other words, if the answer is `24`, you should write "The answer is 24" (as opposed to just showing the code and output).
+  Include the names of anyone you collaborated with at the top of the assignment
+  Homeworks are due at the start of class -- please print the HTML and hand in. Be sure to staple your homework and ensure that it is in an easily readable format. 
+  Labs are not turned in
+  You can delete this header for each assignment


## Question 1

Your answer goes here

```{r}
## Your code goes below
```


## Question 2

Your answer goes here

```{r}
## Your code goes below
```
