---
title: "Lab 2: Hot hand in basketball"
output:
  pdf_document: default
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
author: Michael Lopez, Skidmore College
---



```{r setup, include=FALSE, eval = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Important note 1

First, we'll open RStudio by going to [http://r.skidmore.edu/](http://r.skidmore.edu/). 

# Important note 2

Open a new R Markdown file (File / New File / R Markdown...).  You can create a basic name -- Lab0, for example -- and that'll set you up with a new file ready to go.

# Important note 3

The base RMarkdown file is not particularly useful. Instead, click on the link below:

[https://raw.githubusercontent.com/statsbylopez/FYE_18/master/HWs/HW_lab_base.Rmd](https://raw.githubusercontent.com/statsbylopez/FYE_18/master/HWs/HW_lab_base.Rmd)

Copy and paste the text at the link above into your Markdown document. 

Set the output to HTML mode, and click "Knit HTML" at the top (or Command+Shift+K on Mac).  This should produce a web page with the knitting procedure executing your code blocks.  You can edit this new file to produce your homework submission. You will also print this and hand in for each homework assignment.


##  Lab 3:  Hot hand in basketball

Basketball players who make several baskets in succession are described as 
having a *hot hand*. Fans and players have long believed in the hot hand 
phenomenon, which refutes the assumption that each shot is independent of the 
next. However, [a 1985 paper](http://www.sciencedirect.com/science/article/pii/0010028585900106) by Gilovich, Vallone, and Tversky collected evidence
that contradicted this belief and showed that successive shots are independent 
events. This paper started a great controversy that continues to this day, as you can 
see by Googling *hot hand basketball*.

We do not expect to resolve this controversy today. However, in this lab we'll 
apply one approach to answering questions like this. The goals for this lab are 
to (1) think about the effects of independent and dependent events, (2) learn 
how to simulate shooting streaks in R, and (3) to compare a simulation to actual
data in order to determine if the hot hand phenomenon appears to be real.
To get you started, enter the following in one of the code chunks. 

```{r, eval=FALSE}

library(RCurl)
library(tidyverse)
url <- getURL("https://raw.githubusercontent.com/JunWorks/NBAstat/master/shot.csv")
nba.shot <- read.csv(text = url)
nba.threes <- nba.shot %>% 
  filter(PTS_TYPE == 3) %>% 
  na.omit()
nba.threes %>% head()

nba.threes.sorted <- nba.threes %>% 
  arrange(playerName, GAME_ID, SHOT_NUMBER) %>% 
  group_by(playerName, GAME_ID) %>% 
  mutate(THREE_NUMBER = 1:n()) %>% 
  select(playerName, GAME_ID, THREE_NUMBER, FGM, SHOT_DIST, LOCATION) 

nba.threes.sorted %>% head()
```

1. What is the newly created variable? 

2. The code above filters to only look at three-pointers. What are advantages and disadvantages of this decision?



## Understanding context

One of the major problems in looking at things like the hot-hand is that certain players are better than others. Let's focus on one player only: Nick Young.

```{r, eval = FALSE}
nick.threes <- nba.threes.sorted %>% 
  filter(playerName == "Nick Young") %>% 
  ungroup() %>% 
  mutate(shot.order = 1:n(), 
         shot.abbrev = ifelse(FGM == 1, "H", "M"))
```

First, we make a plot of his makes and misses, all ordered from the beginning of the year to the end of the year. 

```{r, eval = FALSE}
ggplot(data = nick.threes, aes(x = shot.order, y = FGM, colour = FGM)) + 
  geom_point()
```

3. Identify points in the year where it looks like Nick Young was "hot" or where Nick Young was "cold". 

4. Take a look at all of his three-point results using the command `nick.threes$FGM`, which prints out a vector of 0's and 1's. For example, Nick had the following sequence of hits and misses from his first ten shots: 0,1,0,1,0,0,1,0,0,0. See how your conclusion lines up to your answer in the question above. 

## How to define streakiness?

Let's say a player had the following shot run. To make things a bit more obvious, we use hit's (H) and misses (M) to abbreviate shot results

\[ \textrm{H M | M | H H M | M | M | M} \]

Within the nine shot attempts above, there are six streaks, which are separated by a 
"|" above. Their lengths are one, zero, two, zero, zero, zero (in order of 
occurrence).

5. What does it mean to have a streak length of 4, using the definition above? 

Here's a function in R that can calculate the number of streaks for certain runs of shots. 

```{r}
calc_streak <- function(x) {
  if (!is.atomic(x))
    x <- x[, 1]
  if (any(!x %in% c("H","M")))
    stop('Input should only contain hits ("H") and misses ("M")')

    y <- rep(0, length(x))
    y[x == "H"] <- 1
    y <- c(0, y, 0)
    wz <- which(y == 0)
    streak <- diff(wz) - 1

    return(data.frame(length = streak))
}
nick_streak <- calc_streak(nick.threes$shot.abbrev)
head(nick_streak)

ggplot(data = nick_streak, aes(length)) + geom_histogram()
```



5. Describe the distribution of Nick's streak lengths. 
    What was his typical streak length? How long was his longest streak of 
    baskets? 
    

## Compared to What?

We've shown that Nick had some long shooting streaks, but are they long enough 
to support the belief that he had a hot hand? What can we compare them to?

To answer these questions, let's return to the idea of *independence*. Two 
processes are independent if the outcome of one process doesn't effect the outcome 
of the second. If each shot that a player takes is an independent process, 
having made or missed your first shot will not affect the probability that you
will make or miss your second shot.

A shooter with a hot hand will have shots that are *not* independent of one 
another. Specifically, if the shooter makes his first shot, the hot hand model 
says he will have a *higher* probability of making his second shot.

Let's suppose for a moment that the hot hand model is valid for Nick During his
career, the percentage of time Nick makes a 3-pointer (i.e. his shooting 
percentage) is about 41%.

If he makes the first shot and has a hot hand (*not* independent shots), then 
the probability that he makes his second shot would go up to, let's say, 60%. 

As a result of these increased probabilites, you'd expect Nick to have longer 
streaks. Compare this to the skeptical perspective where Nick does *not* have a
hot hand, where each shot is independent of the next. If he hit his first shot,
the probability that he makes the second is still 0.45.

In other words, making the first shot did nothing to effect the probability that
he'd make his second shot. If Nick's shots are independent, then he'd have the 
same probability of hitting every shot regardless of his past shots: 45%.

Now that we've phrased the situation in terms of independent shots, let's return
to the question: how do we tell if Nick's shooting streaks are long enough to 
indicate that he has a hot hand? We can compare his streak lengths to someone
without a hot hand: an independent shooter. 

## Simulating data

Here's a way that we can simulate the three-pointers that Nick took in a season. 

```{r}
n.shots = nrow(nick.threes)
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = n.shots, prob = c(0.41, 0.59), replace = TRUE)
n.shots
sim_basket

nick_streak_sim <- calc_streak(sim_basket)
ggplot(data = nick_streak_sim, aes(length)) + geom_histogram()
```

6. Is Nick Young streaky? Re-run the code above to get several sets of plots. You probably want to do 5-10 iterations. Compare these simulated shot-streaks to the observed Nick Young shot length from above. 


