---
title: "Introduction to R and RStudio using baseball stats"
output:
  pdf_document: default
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
author: Michael Lopez, Skidmore College
---

The movie [Moneyball](http://en.wikipedia.org/wiki/Moneyball_(film)) focuses on
the "quest for the secret of success in baseball". It follows a low-budget team, 
the Oakland Athletics, who believed that underused statistics, such as a player's 
ability to get on base, betterpredict the ability to score runs than typical 
statistics like home runs, RBIs (runs batted in), and batting average. Obtaining 
players who excelled in these underused statistics turned out to be much more 
affordable for the team.

In this lab we'll be looking at data from all 30 Major League Baseball teams and
examining the relationships between in-game statistics. Our primary aim, for today, is a familiarity with R and RStudio, which you'll be using throughout the course both to learn the statistical concepts discussed in class and also to analyze real data and come to informed conclusions.  

To straighten out which is which: R is the name of the programming language itself and RStudio is a convenient interface.

As the labs progress, you are encouraged to explore beyond what the labs dictate;
a willingness to experiment will make you a much better programmer.  Before we 
get to that stage, however, you need to build some basic fluency in R.  Today we
begin with the fundamental building blocks of R and RStudio: the interface, 
reading in data, and basic commands.

The panel in the upper right contains your *workspace* as well as a history of 
the commands that you've previously entered.  Any plots that you generate will 
show up in the panel in the lower right corner.

The panel on the left is where the action happens.  It's called the *console*. 
Everytime you launch RStudio, it will have the same text at the top of the 
console telling you the version of R that you're running.  Below that information
is the *prompt*.  As its name suggests, this prompt is really a request, a 
request for a command.  Initially, interacting with R is all about typing commands
and interpreting the output. These commands and their syntax have evolved over
decades (literally) and now provide what many users feel is a fairly natural way
to access data and organize, describe, and invoke statistical computations.

To get you started, enter the following command at the R prompt (i.e. right after
`>` on the console).  You can either type it in manually or copy and paste it
from this document. This command instructs R to access the OpenIntro website and fetch some data: the `mlb11` data set.

```{r, eval=TRUE}
download.file("http://www.openintro.org/stat/data/mlb11.RData", destfile = "mlb11.RData")
load("mlb11.RData")
```

Let's load up the data for the 2011 season. In addition to runs scored, there are seven traditionally used variables in the data set: at-bats, hits, home runs, batting average, strikeouts, stolen bases, and wins. There are also three newer variables: on-base percentage, slugging percentage, and on-base plus slugging. For the first portion of the analysis we'll consider the seven traditional variables. At the end of the lab, you'll work with the newer variables on your own.

You should see that the workspace area in the upper righthand corner of the RStudio window now lists a data set called `mlb11` that has 30 observations on 12 variables. As you interact with R, you will create a series of objects. Sometimes you load them as we have done here, and sometimes you create them yourself as the byproduct of a computation or some analysis you have performed. Note that because you are accessing data from the web, this command (and the entire assignment) will work in a computer lab, in the library, or in your dorm room; anywhere you have access to the Internet.



## The Data: 2011 baseball statistics

```{r, eval=FALSE}
mlb11
```


What you should see are four columns of numbers, each row representing a 
different team: the first entry in each row is simply the row number (an index 
we can use to access the data from individual years if we want), the second is 
the team, and the remaining columns are team-specific metrics. Use the scrollbar on the right side of the console window to examine the complete data set.

Note that the row numbers in the first column are not part of the MLB data. 
R adds them as part of its printout to help you make visual comparisons. You can
think of them as the index that you see on the left side of a spreadsheet. In 
fact, the comparison to a spreadsheet will generally be helpful. R has stored 
MLB data in a kind of spreadsheet or table called a *data frame*.

You can see the dimensions of this data frame by typing:

```{r, eval=TRUE}
dim(mlb11)
```

This command should output `[1] 30 12`, indicating that there are 30 rows and 12 
columns (we'll get to what the `[1]` means in a bit), just as it says next to 
the object in your workspace. You can see the names of these columns (or 
variables) by typing:

```{r, eval=TRUE}
names(mlb11)
```

You should see that the data frame contains the columns `team`,  `runs`, and 
`at_bats`, etc. At this point, you might notice that many of the commands in R look a 
lot like functions from math class; that is, invoking R commands means supplying
a function with some number of arguments. The `dim` and `names` commands, for 
example, each took a single argument, the name of a data frame. 

One advantage of RStudio is that it comes with a built-in data viewer. Click on
the name `mlb11` in the *Environment* pane (upper right window) that lists 
the objects in your workspace. This will bring up an alternative display of the 
data set in the *Data Viewer* (upper left window). You can close the data viewer
by clicking on the *x* in the upper lefthand corner.


## The tidyverse

There are additional sets of packages called the `tidyverse` that streamline most of the commands that you will need in this course. 

To use the package, you will first need to load it.

```{r, eval=FALSE}
library(tidyverse)
```

Note that you will have to do this *every* time you start a new R session. 

The tidyverse is based on the `%>%` symbol, called a pipe, which takes whatever code is written before the pipe and accordingly does something with it. 

Two useful examples: 

One more super-useful command, `head()`, which shows the first six rows of any data frame.

```{r}
mlb11 %>% head()
```

Related: `tail()'

```{r}
mlb11 %>% head()
```

In the above instances, the `%>%` (the pipe) takes our initial data set and sends it to the next command (either `head` or `tail`)

### Selecting certain variables: select()

Let's start to examine the data a little more closely. We can access the data in
a single column of a data frame separately using a command like

```{r, eval=FALSE}
mlb11 %>% select(runs)
```

This command will only show the number of runs scored each year by each team.

1.  What command would you use to extract just the number of hits for each team? Try it!

To obtain several variables at once, we can use the same `select()` command, with a comma in between variables. 

```{r, eval = FALSE}
mlb11 %>% select (hits, runs, strikeouts)
```

### Creating new variables: `mutate()`

R can do lots of things besides graphs. In fact, R is really just a big calculator. We can type in mathematical expressions like

```{r, eval=FALSE}
210 + 930
```

to see the total number of homeruns and strikeouts by the Texas Rangers, the first row in our data. 

We could do this for every row, but there is a faster way: `mutate()`

```{r, eval=FALSE}
mlb11 <- mlb11 %>% 
  mutate(total.hr.k = homeruns + strikeouts)

mlb11 %>% head()
```

Can you see the new variable called `total.hr.k`?  

One important aspect about naming variables: keep them simple, and avoid funny symbols. The above is named `total.hr.k`, but could have easily been named `total.hr.strikeout` or something like that, or even `total.hr.proflopezrules.strikeout` if you got antsy. A bad variable name? `strikeouts and ks`, or `strikeouts&%ks`, or `ToTaLkS%&sc.` I don't know what that is, but it shouldn't be a variable. 


In addition to simple mathematical operators like subtraction and 
division, you can ask R to make comparisons like greater than, `>`, less than,
`<`, and equality, `==`. For example, we can ask which home run observations were greater than 200.

```{r, eval=FALSE}
mlb11 <- mlb11 %>% 
  mutate(big.power = mlb11$homeruns > 200)
```

This output shows a different kind of data than we have considered so far. In the 
`mlb11` data frame, most of our values are numerical. Here, we've asked R to create *logical* data, data where the values are either `TRUE` or `FALSE`. In general, data analysis will involve many different kinds of data types, and one reason for using R is that it is able to represent and compute with many of them.

2.  Which MLB teams had more than 200 home runs in 2011?


### Finding high's and low's: `arrange()`

The `arrange` command can make identifying the best or worst teams or players quite simple. 

```{r}
mlb11 %>% arrange(runs)
mlb11 %>% arrange(-runs)
```

3. What is the difference between each of the two lines of code above? 

4. Which team had the highest cumulative *number of strikeouts* and homeruns in 2011?


### Graphs in the tidyverse: `ggplot`

R has some powerful functions for making graphics. 

The centerpiece of the `tidyverse` syntax is the use of the *ggplot* package. In this case, `gg` stands for grammar of graphics. 

For example, we can create a simple plot of the number of runs each team scored versus its hits


```{r, eval=FALSE, message=FALSE}
ggplot(mlb11, aes(x = runs, y = hits) + 
         geom_point() 
```

By default, R creates a scatterplot with each x,y pair indicated by an open circle. The plot itself should appear under the *Plots* tab of the lower right panel of RStudio. Notice that the command above again looks like a function, this time with two arguments separated by a comma.  The first argument specifies the data, and the two aesthetics (aes) specify the variables. 

5.  Is there an association between runs scored and hits? How would you describe it? Would you consider this plot a univariate plot or a bivariate plot? 

Now, suppose we want to plot only the total number of runs.

```{r, eval=FALSE}
ggplot(mlb11, aes(x = runs) + 
         geom_histogram() 
       
ggplot(mlb11, aes(x = runs) + 
         geom_boxplot() 
```

6.  Describe the distribution of runs scored among MLB teams in 2011.  What is the median? Would you consider this a univariate plot or a bivariate plot? 


## On your own

1. Arrange the data set by the `stolen_bases` variable. Which teams had the most and fewest stolen bases? 

```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


2. Make a graph of team wins during this season. Is it more appropriate to use a boxplot or a scatter plot? Is the distribution of wins skewed left, right, or symmetric?

```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


3. Describe the center, shape, and spread of the `stolen_bases` variable, using an appropriate plot.


```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


4. How can you change the x and y labels on your plots? How can you add a title? Use [google](www.google.com) to guide you, and update your plot in the question above. One trick: include `ggplot` in your google search.

```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


5. Using visual evidence, find the variable that you think seems to boast the strongest association to `runs`. Consider any continuous variables between columns 3 (`at_bats`) and 12 (`new_obs`).

```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


6. A coach is interested in the link between `stolen_bases` and `runs`. Show the coach a scatter plot, and describe the association. As you make the plot, think carefully about which of these two variables is the explanatory variable (and which is the response).


```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


7. Create a new variable for whether or not a team won 90 games or more. You can call this variable whatever name you want. How many teams fit won 90 games or more? Use your new variable to find out.


```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```

8. Knit your document, and cross-check to make sure all of your answers and text are written in the appropriate spots. 



<div id="license">
Portions of this lab were adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel from a lab written by Mark Hansen of UCLA Statistics, a product of OpenIntro that is released under a 
[Creative Commons License](http://creativecommons.org/licenses/by-sa/3.0)
</div>


