---
title: "Introduction to R and RStudio using baseball stats"
output:
  pdf_document: default
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
author: Michael Lopez, Skidmore College
---




## Summaries and tables

A good first step in any analysis is to distill all of that information into a few summary statistics and graphics.  As a simple example, the function `summary` returns a numerical 
summary: minimum, first quartile, median, mean, second quartile, and maximum. 
For `runs` this is

```{r, eval=FALSE}
summary(mlb11$runs)
```

If you wanted to compute the interquartile range for team runs, you could look at the output from the summary command above and then enter

```{r, eval=FALSE}
734 - 629
```

R also has built-in functions to compute summary statistics one by one.  For 
instance, to calculate the mean, median, and variance of `runs`, type 

```{r, eval=FALSE, message=FALSE}
var(~runs, data=mlb11)
median(~runs, data=mlb11)
```

The `mosaic` command `favstats`, allows us to compute all of this information (and more) at once. 

```{r, eval=FALSE}
favstats(~runs, data=mlb11)
```

While it makes sense to describe a quantitative variable like `runs` in terms of these statistics, what about categorical data?  We would instead consider the sample frequency or relative frequency distribution.  The function `tally` does this for you by counting the number of times each kind of response was given. For example, to see the number of teams who hit at least 200 homeruns, type

```{r, eval=FALSE}
mlb11$HighHR <- mlb11$homeruns >= 200
mlb11$HighHR
tally(~ HighHR, data=mlb11)
```

or instead look at the relative frequency distribution by typing

```{r, eval=FALSE}
tally(~ HighHR, data=mlb11, format="proportion")
```

In each of the above steps, we make use of a new variable, `HighHR` (the contents of which 
we can see by typing `HighHR` into the console) and then used it in as the input for `tally`. The special symbol `<-` performs an *assignment*, taking the output of one line of code and saving it into an object in your workspace.  This is another important idea that we'll return to later.


Notice how R automatically divides all entries in the table by 30 in the second command. This is similar to something we observed earlier; when we multiplied or divided a vector with a number, R applied that action across entries in the vectors. As we see above, this also works for tables. Next, we make a bar chart of the entries in the table by putting the table inside the`barchart` command.

```{r, eval=FALSE}
barchart(tally(~HighHR, data=mlb11, margins=FALSE), horizontal=FALSE)
```

Notice what we've done here! We've computed the table of and 
then immediately applied the graphical function, `barchart`. This is an 
important idea: R commands can be nested. You could also break this into two 
steps by typing the following:

```{r, eval=FALSE}
HR <- tally(~ HighHR, data=mlb11, margins=FALSE)
barchart(HR, horizontal=FALSE)
```


6.  Create a numerical summary for `stolen_bases` and `wins`, and compute the interquartile range for each. 


7.  Compute the relative frequency distribution for teams with at least 90 `wins`. How many teams reached 90 wins?  


The `tally` command can be used to tabulate any number of variables that you 
provide.  

8.  What is shown in the following table?  How many teams won 90 or more games and hit at least 200 home runs? 


```{r table-smoke-gender, eval=FALSE}
mlb11$HighWins <- mlb11$wins >=90
tally(HighWins ~ HighHR, data=mlb11, format="count")
```


## How R thinks about data

We mentioned that R stores data in data frames, which you might think of as a 
type of spreadsheet. If we want to access a 
subset of the full data frame, we can use row-and-column notation. For example,
to see the sixth variable of the 7<sup>th</sup> respondent, use the format

```{r, eval=FALSE}
mlb11[7,6]
```

which means we want the element of our data set that is in the 7<sup>th</sup> 
row (meaning the 7<sup>th</sup> team) and the 6<sup>th</sup> 
column (in this case, batting average). We know that `bat_avg` is the 6<sup>th</sup> variable because it is the 6<sup>th</sup> entry in the list of variable names

```{r, eval=FALSE}
names(mlb11)
```

To see the `bat_avg` for the first 10 respondents we can type

```{r, eval=FALSE}
mlb11[1:10,6]
```

In this expression, we have asked just for rows in the range 1 through 10.  R 
uses the `:` to create a range of values, so 1:10 expands to 1, 2, 3, 4, 5, 6, 
7, 8, 9, 10. You can see this by entering

```{r, eval=FALSE}
1:10
```

Finally, if we want all of the data for the first 10 respondents, type

```{r, eval=FALSE}
mlb11[1:10,]
```

By leaving out an index or a range (we didn't type anything between the comma 
and the square bracket), we get all the columns. When starting out in R, this is
a bit counterintuitive. As a rule, we omit the column number to see all columns 
in a data frame. Similarly, if we leave out an index or range for the rows, we 
would access all the observations, not just the 7<sup>th</sup>, or rows 1 
through 10. Try the following to see the batting averages for all 30 teams.

```{r, eval=FALSE}
mlb11[,6]
```

There's an easier way, which you will recall from earlier

```{r, eval=FALSE}
mlb11$bat_avg
```

The dollar-sign tells R to look in data frame `mlb11` for the column called 
`bat_avg`.  Since that's a single vector, we can subset it with just a single 
index inside square brackets.  We see the batting average for the 7<sup>th</sup> 
respondent by typing

```{r, eval=FALSE}
mlb11$bat_avg[7]
```


## Exercises. 


9.  Describe the center, shape, and spread of the `stolen_bases` variable, using an appropriate plot and the appropriate metrics.


```{r}  
### Enter code here.
### Note that any comments in code after the `#` do not get read by R


```


10.  A coach is interested in the link between `stolen_bases` and `runs`. Show the coach a scatter plot, and describe the association. As you make the plot, think carefully about which of these two variables is the explanatory variable (and which is the response).


```{r}
### Enter code here.


```

11.   How can you change the x and y labels on your plots? How can you add a title? Use [google](www.google.com) to guide you, and update your plot in Question 10.


```{r}
### Enter code here.


```


12.   Using visual evidence, find the variable that you think seems to boast the strongest association to `runs`. Consider any continuous variables between columns 3 (`at_bats`) and 12 (`new_obs`).


```{r}
### Enter code here. 


```

13.   What is the variance in the number of strikeouts for each team during the 2011 season? 

```{r}
### Enter code here.

```


14.   Make a new variable, `high_BA`, to represent teams that hit for a batting average of 0.270 or more. How many teams fit into this group?

```{r}
### Enter code here

```

15.   You can use favstats to get the summary statistics within each `high_BA` group, and `bwplot()` to make boxplots. For example, what appears to be the link between `homeruns` and `high_BA`? Do teams with higher batting averages tend to hit more home runs?

```{r, eval = FALSE}
favstats(homeruns ~ high_BA, data = mlb11)
densityplot(~ homeruns, groups = high_BA, data = mlb11, auto.key = TRUE)
```

16.   Repeat the code above, only using `stolen bases` instead of `homeruns`. Does there appear to be a link between `stolen bases` and `high_BA`? 

```{r, eval = FALSE}
### Enter code here. 


```



That was a short introduction to R and RStudio, but we will provide you with more functions and a more complete sense of the language as the course progresses. Feel free to browse around the websites for [R](http://www.r-project.org) and [RStudio](http://rstudio.org) if you're interested in learning more.



<div id="license">
Portions of this lab were adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel from a lab written by Mark Hansen of UCLA Statistics, a product of OpenIntro that is released under a 
[Creative Commons License](http://creativecommons.org/licenses/by-sa/3.0)
</div>


